## Stage 2
For this stage, I have created an RStudio Server for all of you to use. You can access this server:

rstudio.sullivanstatistics.com
For sign in:

**Username**: rsknkr
**Password***: everyone has the same password: password

For the setup of this, I have created a database called titanic. The goal will be to use this lab to work on the titanic dataset. The code below will let you connect to the database. Create a Markdown with the following code and work through the lab.
```{r}
library(RMySQL)
con <- dbConnect(MySQL(),
                 user = 'trial',
                 password = 'password',
                 host = '127.0.0.1',
                 dbname='titanic')

```
We can run any of these commands as follow:
```{r}
query <- "show tables;" # This will be your SQL command
dbGetQuery(con, query) # This sends the command to the SQL server
```  
1. What tables are in this database?
**There are tables names of which start with "titanic" (all tables created by students ever enrolled in PHP2560, I suppose) and also tables "attempt" and "table_borg"**

2. Load the titanic data:
```{r}
titanic <- read.table("/home/php2560/titanic.csv", header=TRUE, sep=",")
```
3. Create a table in MySQL from this data. Name it titanic_raskina:

```{r, eval=FALSE}
dbWriteTable(con, "titanic_raskina", titanic)
```

Now what tables are in this database?

```{r}
con2 <- dbConnect(MySQL(),
                 user = 'trial',
                 password = 'password',
                 host = '127.0.0.1',
                 dbname='titanic')
query <- "show tables;" # This will be your SQL command
dbGetQuery(con2, query)
```
**There's one more table (titanic_raskina) in this database now**

### Basic usage of commands

We will start to use a list of commands that would be commonly used in MySQL:

These are:

- show tables;
- describe titanic_yourlastname;
- select name, pclass, age, sex from titanic limit 10;

Try these commands and test this out. 


### R and MySQL

5. Run the following SQL command. Before doing so describe what this is asking
```{r}
# This command selects columns pclass and survived from my table and also creates a column avg_age, where the average age of those who survived or didn't survive in each class is shown

query <- "SELECT pclass, survived, avg(age) AS avg_age FROM titanic_raskina   
   GROUP BY pclass, survived"
dbGetQuery(con2, query)

```


6.  Run the following SQL command. Before doing so describe what this is asking
```{r}
# This command selects columns pclass and survived from my table and also creates a column avg_age, where the average age of only those who survived in each class is shown

query <- "select pclass, survived, avg(age) as avg_age from titanic_raskina
   where survived=1
   group by pclass, survived"
dbGetQuery(con2, query)

```


7. Run the following SQL command. Before doing so describe what this is asking
```{r}
# This command selects columns name, pclass, survived and age from my table for those passengers whose full name matches the pattern '^.sen.+' (contains parts which start with ".sen", where "." is  single character, and have at least 1 character after this pattern)

query <- "select name, pclass, survived, age from titanic_raskina 
   where name regexp '^.sen.+'"
dbGetQuery(con2, query)

```

8. Run the following SQL command. Before doing so describe what this is asking
```{r}
# This command selects columns name, pclass, survived and age from my table for those passengers whose full name matches the pattern 'Jakob.*$' (ends with "Jakob" and may or may not have characters after this pattern)

query <- "select name, pclass, survived, age from titanic_raskina 
   where name regexp 'Jakob.*$'"
dbGetQuery(con2, query)

```

9.  Run the following SQL commands. What is different about them? What do they return?
```{r}
# This command selects columns name, pclass, survived and age from my table for those passengers whose full name matches the pattern 'Ivar.*$' (ends with "Ivar" and may or may not have characters after this pattern)


query <- "select name, pclass, survived, age from titanic_raskina 
   where name regexp 'Ivar.*$'"
dbGetQuery(con2, query)
```
   
```{r}
# This command selects columns name, pclass, survived and age from my table for those passengers whose first name matches the pattern 'Ivar.*$' (ends with "Ivar" and may or may not have characters after this pattern)

# The result will not contain those rows from the previous command where "Ivar.*$" is located before the comma (in other words, where this pattern is a part of a last name)

query <- "select name, pclass, survived, age from titanic_raskina 
   where name regexp ',.*Ivar.*$'"
dbGetQuery(con2, query)
```


10. We can also plot data from this:
```{r}
myQuery <- "select pclass, avg(fare) as avg_fare from titanic_raskina
              group by pclass"
myData <- dbGetQuery(con, myQuery)
library(ggplot2)
ggplot(myData, aes(pclass, avg_fare)) + geom_bar(stat="identity")
```



### More challenging

11. Create a data frame with:
- `pclass`
- `sex`
- `age`
- `fare`
- `survived`
- `parch`
Only for people who did not survive. 

```{r}
query <- "select pclass, sex, age, fare, survived, parch from titanic_raskina where survived = 0"
dbGetQuery(con2, query)

```

12. Create a graph of the average survival over the different classes. Make two distinct lines for male and female passengers. Create this plot using ggplot. 
```{r}
myQuery <- "select pclass, sex, avg(survived) as average_survival from titanic_raskina
              group by pclass, sex"
myData <- dbGetQuery(con2, myQuery)

library(ggplot2)
ggplot(myData, aes(x=pclass, y=average_survival, color=sex)) + scale_x_discrete(limits=c("1","2","3")) + geom_line()

```



