# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50.

```{r}
##Let's make a loop to print all integers from 1 to 50
for (i in 1:50) {print(i)}
```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
##Now let's add all integers from 0 to 1000
sumAll <- 0
for (i in 0:1000) {sumAll <- sumAll + i}
```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
##Now let's add all EVEN integers from 0 to 1000
sumEven <- 0
for (i in seq(from = 0, to = 1000, by = 2)) {sumEven <- sumEven + i}
```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
##No loop this time
sumAll <- sum(seq(0,1000))
sumEven <- sum(seq(0, 1000, 2))
```


### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
survey
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean <- survey
question <- c("q1", "q2", "q3", "q4", "q5")
for (q in question) {
  for (r in 1:5) {
    if (survey.clean[r,q] < 1 | survey.clean[r,q] > 5) {survey.clean[r,q] = NA}
  }
}
survey.clean

```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
survey.clean <- survey
survey.clean$invalid.answers <- c(0,0,0,0,0,0)

question <- c("q1", "q2", "q3", "q4", "q5")

#double loop: for columns (q) and then for rows (r)
for (q in question) {
  for (r in 1:5) {
    if (survey.clean[r,q] < 1 | survey.clean[r,q] > 5) {
      survey.clean[r,q] <- NA
      survey.clean[r,"invalid.answers"] <- survey.clean[r,"invalid.answers"] + 1
      
            }
  }
}
survey.clean

```