# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
# function takes two character vectors as parameters
# substr extracts elements from the "string", #1 to #(length of the string passed as "prefix"") 
# if the substring is equal to "prefix", function returns TRUE, otherwise - FALSE

str_equality <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
```

```{r}
# function takes a vector as a parameter
# returns the vector without the last element; if there was only one element, returns NULL 
cut_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
```

```{r}
# function takes two vectors as parameters
# the output of the function is a vector of the same length as vector x that consists of replicated values of vector y
repl_vec <- function(x, y) {
  rep(y, length.out = length(x))
}

```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
